An efficient composition of bidirectional programs by tupling and lazy updates

### 2019-11-25 paper/2.minbigul.tex
Update 2.minbigul.tex

MinBiGUL, our target language in this paper, is a subset of BiGUL [?] which is a simple yet powerful putback-based bidirectional language. BiGUL supports two transformations: a forward transformation $get$ producing a view from a source and a backward transformation $put$ taking a source and a modified view to produce an updated source. Intuitively, if we have a BiGUL program $bx$, these two transformations are following functions:\\
    $\tab \getbxinline{bx} : s \to v,
    \tab \putbxinline{bx} : s * v \to v$

BiGUL is well-behaved since two functions $\putbxinline{bx}$ and $\getbxinline{bx}$ satisfy the round-trip laws as follows:\\
    $\tab \putbx{bx}{s}{(\getbx{bx}{s})} = s \qquad [\textsc{GetPut}]\\
    \tab \getbx{bx}{(\putbx{bx}{s}{v})} = v \qquad [\textsc{PutGet}]$

The \textsc{GetPut} law means that if there is no change to the view, there should be no change to the source. The \textsc{PutGet} law means that we can recover the modified view by applying the forward transformation to the updated source.

MinBiGUL inherited from BiGUL also supports transformations $put$ and $get$ which are satisfy two above laws. In addition, we restrict adaptive cases of BiGUL on minBiGUL. Then $put$ and $get$ satify one more law, \textsc{PutPut}, like the following:\\
    $\tab \putbx{bx}{(\putbx{bx}{s}{v'})}{v} = \putbx{bx}{s}{v} \qquad [\textsc{PutPut}]$

The \textsc{PutPut} law means that a source update should overwrite the effect of previous source updates. Due to the satisfaction of three laws, \textsc{GetPut}, \textsc{PutGet} and \textsc{PutPut}, minBiGUL is very well-behaved.

@@

For source/view rearrangement, BiGUL uses just one lambda expression to express how to deconstruct as well as reconstruct data. It is a kind of bijection. However, to be able to implement it in OCaml, the environment used for developing minBiGUL and solutions in the paper, we need give two functions which one is the inverse of the other. In the above syntax, $f_2 = f_1^{-1}$ and $g_2 = g_1^{-1}$.

To help make demonstration more direct, we provide the following alternatives representation: 
    $\quad Prod \ bx_1 \ bx_2 \equiv bx_1 \times bx_2,
    \quad Compose \ bx_1 \ bx_2 \equiv bx_1 \circ bx_2$

In general, $\circ$ has a higher priority than $\times$. Since only consider a product of two minBiGUL programs is considered, it is unnecessary to define the associativity precedence of $\times$. The one of $\circ$ may be left or right, however we do not declare the default.

@@

% Next, let's take a look at some examples to better understand about minBiGUL. We start with quite obivious things as follows:

>>>>

MinBiGUL, our target language in this paper, is a subset of BiGUL which is a simple yet powerful putback-based bidirectional language. BiGUL supports two transformations: a forward transformation $get$ producing a view from a source and a backward transformation $put$ taking a source and a modified view to produce an updated source. Intuitively, if we have a BiGUL program $bx$, these two transformations are following functions:\\
    $\tab \getbxinline{bx} : s \to v,
    \tab \putbxinline{bx} : s * v \to v$

BiGUL is well-behaved [?] since two functions $\putbxinline{bx}$ and $\getbxinline{bx}$ satisfy the round-trip laws as follows:\\
    $\tab \putbx{bx}{s}{(\getbx{bx}{s})} = s \qquad [\textsc{GetPut}]\\
    \tab \getbx{bx}{(\putbx{bx}{s}{v})} = v \qquad [\textsc{PutGet}]$

The \textsc{GetPut} [?] law means that if there is no change to the view, there should be no change to the source. The \textsc{PutGet} [?] law means that we can recover the modified view by applying the forward transformation to the updated source.

MinBiGUL inherited from BiGUL also supports transformations $put$ and $get$ which are satisfy two above laws. In addition, we restrict adaptive cases of BiGUL on minBiGUL. Then $put$ and $get$ satisfy one more law, \textsc{PutPut} [?], like the following:\\
    $\tab \putbx{bx}{(\putbx{bx}{s}{v'})}{v} = \putbx{bx}{s}{v} \qquad [\textsc{PutPut}]$

The \textsc{PutPut} law means that a source update should overwrite the effect of previous source updates. Due to the satisfaction of three laws, \textsc{GetPut}, \textsc{PutGet} and \textsc{PutPut}, minBiGUL is very well-behaved [?].

@@

For source/view rearrangement, BiGUL use just one lambda expression to express how to deconstruct as well as reconstruct data. It is a kind of bijection. However, to be able to implement it in OCaml, the environment used for developing minBiGUL and solutions in the paper, we need give two functions which one is the inverse of the other. In the above syntax, $f_2 = f_1^{-1}$ and $g_2 = g_1^{-1}$.

To help make demonstration more direct, we provide the following alternatives representation: $Prod \ bx_1 \ bx_2 \equiv bx_1 \times bx_2, \ Compose \ bx_1 \ bx_2 \equiv bx_1 \circ bx_2$. In general, $\circ$ has a higher priority than $\times$. Their associativity precedence can be either left or right or mixture, but are not set by default. We need to explicitly write programs that use these operators.

@@

% Next, let's take a look at some examples to better understand about minBiGUL. We start with quite obvious things as follows:
